### 2018-05-15 boneyard.tex
saving progress half way through the event segmentation results section

Another interesting feature of the recall matrices is the off-diagonal structure, which represents the use of language that generalizes to more than one event. To characterize this aspect of the recall structure, we averaged the rows of the recall matrices according to the k events computed in the analysis described above.  This reduced the recall matrices from \# of sentences to k rows. Then, we computed k by k correlation matrices, representing the similarity of each event to each other event. To visualize these recall `networks', we embedded the k by topics matrices into a 2-dimensional space using Multidimensional Scaling (MDS), where each dot represents an event and the color of the dot represents the event index.  Then, we connected each event with a line where the color and line weight was proportional to the strength of the correlation between the events. These network plots can be seen in Fig X. To hypothesized that stronger correlations between recall events would

\subsubsection{Temporally aligning the stimulus and recall}
While most participants were able to recall most of the scenes in order, the rate of verbal description varied by participant and timepoint (Supp Fig X). Because of this participant-dependent non-linear relationship between stimulus and recall time, averaging the video-recall correlation plots across participants attenuates the robust within-participant effects. While further exploration of individual differences in recall rates is interesting, our primary goal here was to develop a model capable of accurately characterizing the content recovered during recall and explore how that varied across participants.  To that end, we leveraged dynamic time warping (DTW, REF) to temporally align the stimulus and recall matrices.  The algorithm returns a path of indices that bring the two timeseries into maximal temporal alignment (within constraints, see Methods for details).  For each participant, we computed the path and then warped the video model and recall model according to it. The length of the warp path varied by participant, and so the length (number of timepoints) of the warped video/recall matrices also varied by participant (range: X-X). To account for these differences, we resampled the timeseries back to the length of the stimulus (1976 time samples).

\subsection{fMRI analyses}
Seventeen participants watched the first 50 minutes of Episode 1 of BBC's Sherlock. The video was split into two parts of approximately equal length (946 and 1030 TRs). All data were preprocessed and transformed to 3mm MNI space as described in the paper. Data were zscored across time at every voxel. 6mm smoothing was applied.
Files are cropped so that all video-viewing data are aligned across participants, and all recall data are aligned to the scene timestamps below. The cropping includes a constant 3-TR (4.5 sec) shift to correct for hemodynamic lag.

\subsubsection{Univariate subsequent memory analysis}
To characterize brain regions involved in memory encoding, we conducted an analysis inspired by prior studies investigating the  `subsequent memory effect' (Paller and Wagner).  For each participant, we computed the correlation between topic vectors at each timepoint of the video model and recall model (they were temporally aligned with DTW and resampled to 1976 timesamples). The result was a 1 x 1976 vector of correlation values representing how similar each moment of the video model was to each moment of the recall model. Then, to compute the average correlation, we Fisher's z transformed the correlation vectors, averaged them together, and then inverse Fisher's z transformed them.  Finally, we mean centered the vector.  This 1 x 1976 vector represents memorable/unmemorable parts of the video, and be thought of as a `continuous' extension of a subsequent memory regressor.  We convolved the timeseries with a double gamma function (PARAMS) to account for the hemodynamic response, and used it as a model for each voxel's activity timeseries (using FSL software).  Finally, to aggregate across participants, we performed a mixed effects (ordinary least squares) analysis, thresholding at p<.001 and cluster correcting a p<.05.

\subsubsection{Multivariate searchlight analyses}
Our multivariate analyses were designed to capture brain regions whose timepoint-by-timepoint correlational structure mirrors the correlational structure of the video and recall topic models during video viewing. To that end, we conducted a searchlight analysis (5x5x5 voxel cube) where for each cube, we correlated the model timepoint-by-timepoint correlation matrix with the neural correlation matrix. To aggregate across participants, we computed Fisher's z transformed the correlations and then averaged.  To assess significance, we recomputed this group analysis 100 times, but randomly phase shifted the model (by \# of timepoints - 1) by the same amount for each participant but different amounts for each permutation to build a null distribution of correlation values. Finally, we thresholded the group averaged correlation maps where the `real' correlation value for a given voxel exceeded the 95th percentile of the null distribution.

We first performed this analysis using the unwarped video model. Thus, the same model was applied to each participant and each searchlight cube. We then performed this analysis on the participant-wise recall models. To correct for non-linearities between the video model and recall models, for each participant we used DTW to temporally align the matrices. The algorithm recovers a path of coordinates that would bring the video and recall model in maximal temporal alignment. We used this path to warp the fMRI data and the recall model into alignment (separately for each participant). Then, we performed the exact same analysis as described above, using a searchlight to correlate the timepoint-by-timepoint recall and neural correlation matrices. The statistics for both analyses were computed as described above.

\subsubsection{Accounting for nonlinearities between video and recall models}
For many participants, the rate of recall changed over time (see Supp Fig X for an example). This becomes problematic if averaging the video/recall correlation matrices because the different recall rates distort the group average matrix. To correct for differences in recall rates across participants, we used dynamic time warping (DTW).  Briefly, DTW finds the path that maximizes the temporal alignment between two timeseries (REF). For each participant, we used DTW to temporally align the video model and the participant-specific recall model. Because the path is different for each participant, the length of the warped matrices varied across participants (range: X to X). To account for this, we resampled the matrices back to their original shape (1976 x 100).